Reasoning About a Machine with Local Capabilities

### 2018-05-14 toplas/paper.tex
Figure placement.
Typos and minor changes

\begin{figure}[htb]
  \begin{tikzpicture}[scale=.7, every node={scale=.7}]
    \draw[fill=gray!20,draw=none] (0,4) rectangle (6,3.5);
    \draw[fill=gray!20,draw=none] (0,0.5) rectangle (6,1);
    \draw[draw=gray!80] (0,0.5) -- (0,4);
    \draw[draw=gray!80] (6,0.5) -- (6,4);
    \draw (0,1) rectangle (6,1.5) node[pos=.5] {\footnotesize previous stack pointer};
    \draw (0,1.5) rectangle (6,2) node[pos=.5] {\footnotesize previous program counter};
    \draw (0,2) rectangle (6,3.5) node[pos=.5] {\footnotesize restore instructions};
    \draw (-3,3.25) node (rp) {return pointer};
    \draw[->] (rp.east) -- node[above] {$\entry$}(-.2,3.25);
  \end{tikzpicture}

@@

\begin{toplas}
  In the future world relations, we must remember all region names in order to
  keep them extensional which is why we need to use masks. It does, however, not
  matter what kind of region is used to mask as it is the region name we must
  retain. This is why we allow $\revoked$ regions to be replaced by other
  regions in the two future world relations. This may seem unnecessary as the
  future world relation allows new regions to be added, but in private future
  worlds, for technical reasons, we need to be able to reinstate old regions
  with the same region names as they had in a past world. Further, we do not
  want public future worlds to be able to prevent reinstation in private future
  worlds. Private future worlds cannot revoke $\perma$ regions, so we cannot
  allow public future worlds to us $\perma$ regions as a mask for a revoked
  region.
\end{toplas}

@@

\begin{figure}[htb]
  \centering
  \begin{align*}
   \readCond{}(\gl)(W) &=  \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{array}{l}
       \exists r \in \var{localityReg}(g,W) \ldotp \\
\quad   \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{array}
    \right\}\\
   \writeCond{}(\iota,\gl)(W) &=  \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & W(r) \text{ is address-stratified and} \\[-0.2cm]
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
   \execCond{}(\gl)(W) &= 
    \left\{
      \npairP{(\var{P}, \start,\addrend)}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start',\addrend'] \subseteq [\start,\addrend], \perm \in \var{P} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start',\addrend',\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
   \entryCond{}(\gl)(W) &= 
    \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{array}{l}
      \forall n' < n \ldotp \forall W' \future W \ldotp\\
      \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{array}
    \right\} \\
   \text{where } \gl = \local \Rightarrow \future = \futurewk \text{ and } \gl = \glob \Rightarrow \future = \futurestr \span
   \end{align*}
\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
\begin{figure}[htb]
  \begin{align*}
  H^\nwl_A,  H^\pwl_A :{} & \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
  \iota^\pwl_A, \iota^\nwl_A :{} & \Regions \\
  \iota^\pwl_A \defeq{} & (\temp,1,=,=,H^\pwl_A) \\
  H^\pwl_A \; s \; \hat{W} \defeq{} & \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = A \land \\
      &\forall \addr \in A \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
    \end{aligned}
        \right\} \union \{\npair[0]{\ms} \} \\
  \iota^\nwl_A \defeq{} & (\temp,1,=,=,H^\nwl_A) \\
  \iota^{\nwl,p}_A \defeq{} & (\perma,1,=,=,H^\nwl_A)\\
  H^\nwl_A \; s \;\hat{W} \defeq{} & \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = A \land \\
      &\forall \addr \in A \ldotp \\
      &\quad \nonlocal{\ms(\addr)} \land\\ 
      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
    \end{aligned}
        \right\} \union \{\npair[0]{\ms} \}
\end{align*}

@@

\begin{toplas}
  In \sectionname~\ref{sec:stack-and-return-pointer}, we require capabilities with write-local permission to be local.
  This would indicate that our logical relation should imply the same, namely that capabilities with write-local permission are local:
  \begin{lemma}[Stack capabilities are local]~
    \label{lem:wl-local}
    \begin{itemize}
    \item If $\memSat{\ms}{W}$ and $\npair{\stdcap} \in \stdvr(W)$ and $\start
      \leq \addrend$ and $\perm \in \{\rwlx,\rwl\}$, then $\gl = \local$.
    \end{itemize}
  \end{lemma}
  In our definition of worlds, nothing prevents a world from having regions that are overlapping.
  This is an issue for the $\stdvr$-relation as it allows different permission-based requirements to be satisfied by different regions.
  In practice, it is not an issue as we will always have a memory satisfaction assumption which doubles as a well-formedness as it prevents the regions from overlapping.
  The memory satisfaction assumption in Lemma~\ref{lem:wl-local} is there to ensure a well-formed world.
\end{toplas}

\subsection{Safety of the Capability Machine}
With the logical relation defined, we can now state the fundamental theorem of
our logical relation: a strong theorem that formalizes the guarantees offered by
the capability machine. Essentially, it says a capability that only grants safe
authority is capability safe as a program counter.
\begin{theorem}[Fundamental Theorem]
  \label{thm:ftlr}
  If one of the following holds:
  \begin{align*}
      & \bullet
        \begin{aligned}[t]
        &\perm = \exec \text{ and }\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwx \text{ and } \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwlx \text{ and }\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
      \end{aligned}
  \end{align*}
  then $\npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)$
\end{theorem}
\begin{toplas}
  \begin{proof}[Proof sketch]
    Induction over $n$. By definition of $\stder(W)$, show
    $\npair[n']{(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms)}
    \in \observations(W)$ assuming $n' \leq n$, $\npair[n']{\reg} \in \stdrr(W)$, and
    $\heapSat[\ms]{n'}{W}$. By definition of $\observations$ let $\ms_f$,
    $\mem'$, and $i \leq n'$ be given and for
    $\Phi = (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms \uplus
    \ms_f)$ assume $\Phi \step[i]
    (\halted,\mem')$ and show there exists $W' \futurestr W$ that part of
    $\mem'$ satisfies. First observe that $i \neq 0$ as $\Phi$ is a non-halted
    configuration, so $\Phi$ takes at least one step, i.e.\ $\Phi \step
    \Phi' \step[i-1] (\halted,\mem')$.

@@

In addition to these syntactic requirements, we also specify standard
expectations for how malloc behaves. Intuitively, we require that when malloc is
invoked with a length argument, then it will return a capability for a fresh
piece of memory of that size. It should be fresh in the sense that malloc has
not already given out a capability for any part of that memory before and will
not do so in the future. Also, when invoked with a nonsensical length argument
such as a negative integer or a capability, malloc should simply fail. However,
formulating these requirements is harder than one might expect. The problem is
that a realistic implementation of malloc needs to rely on internal state that
changes after every invocation, and relies on invariants on that state. We can
only expect that malloc behaves according to its specification, if its internal
state satisfies the current state of the invariants in an executing system. To
express this, we will use the semantic model defined in
Section~\ref{sec:logical-relation}.

@@

\begin{figure}[htb]
  \begin{toplas}
    \centering
\begin{lstlisting}
// push private registers to the stack
  push $r_{\var{priv},1}$
  ...
  push $r_{\var{priv},n}$
// push restoration code to the stack
  push $\encode(i_1)$
  push $\encode(i_2)$
  push $\encode(i_3)$
  push $\encode(i_4)$
// push old pc to the stack
  move r_t1 pc
  lea r_t1 $\var{ofc}$ 
  push r_t1
// push stack pointer to the stack
  push r_stk
// set up protected return pointer
  move r_0 r_stk
  lea r_0 $\var{ofc_\var{rec}}$ // -5 is the offset to the first instruction of
the recovery code
  restrict r_0 $\encodePermPair((\local,\entry))$
// restrict stack capability to unused part
  geta r_t1 r_stk
  plus r_t1 r_t1 1
  getb r_t2 r_stk
  subseg r_stk r_t1 r_t2
// clear unused part of the stack
  mclear r_stk
// clear non-argument registers
  rclear $R$
  jmp r
after:
// pop the restore code
  pop r_t1
  pop r_t1
  pop r_t1
  pop r_t1
// pop the private state into approriate registers
  pop $r_{\var{priv},1}$
  ...
  pop $r_{\var{priv},n}$
\end{lstlisting}
    \caption{Implementation of \texttt{scall} $r(r_{\var{args},1},\dots,
      r_{\var{args},m},r_{\var{priv},1},\dots, r_{\var{priv},n})$. The
      restoration code is given in \figurename~\ref{fig:scall-rest-code}. The
      variable $\var{ofc}$ is the offset to the label after, and
      $\var{ofc_\var{rec}} = -5$ which is the offset to the first instruction of
      the activation record. The set $R = \RegName \setminus
      \{\texttt{pc},\texttt{r\_stk},\texttt{r\_0},r,r_{\var{args},1},\dots,r_{\var{args},m}\}$.}
    \label{fig:scall-impl}
  \end{toplas}

\end{figure}
\begin{figure}[htb]
  \begin{toplas}
    \centering
\begin{lstlisting}
 $i_1$ = move r_t1 pc
 $i_2$ = lea r_t1 $\var{ofc}$
 $i_3$ = load r_stk r_t1
 $i_4$ = pop pc
\end{lstlisting}
    \caption{The restoration code used in \texttt{scall}. The variable
      $\var{ofc}$ is 5 which is the offset to the address where the old stack
      pointer is stored on the stack.}
    \label{fig:scall-rest-code}
  \end{toplas}
\end{figure}

@@

% Anti-reduction lemma
Our capability machine only allows the final memory of an execution to be
observed, so naturally the correctness lemmas we prove in
\sectionname~\ref{sec:examples} are statements about the memory in the halted
configuration. In order to prove something about some part of the final memory,
we establish an invariant for it as a region in a world and show that the
initial configuration is in the $\observations$-relation w.r.t.\ that world. The
$\observations$-relation says that if the initial configuration halts
successfully, then the memory in the halting state must still respect a private
future world of the initial world. It is, however, bothersome and error prone to
try to argue about the entire execution in one go. Instead we want to reason
modularly, in the sense that we only want to reason about parts of the execution
at a time. To this end, we prove an anti-reduction lemma that essentially says
that if we can show that an initial configuration, $\Phi$, steps to a
configuration $\Phi'$ and that $\Phi'$ is in the $\observations$-relation, then
$\Phi$ is also in the observation relation.
\begin{lemma}[Anti-reduction for $\observations$]
\label{lem:anti-red-obs}
  \begin{align*}
    & \forall n, n',i,\reg, \reg', \ms, \ms', \ms_r, W, W' \ldotp\\
    & \quad n' \geq n - i \land W' \futurestr W \land \\
    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \\
    & \quad \npair[n']{(\reg',\ms')} \in \observations(W') \\
    & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
  \end{align*}
\end{lemma}

% FTLR
The anti-reduction works well when we reason about the execution of known code,
because we can execute it. When we want to reason about unknown code, the
anti-reduction lemma does not apply because we do not know what instructions are
being executed. This is where the FTLR (Theorem~\ref{thm:ftlr}) comes into play.
As a reminder, the FTLR says that if a capability only has access to safe values
with respect to a world, then it is safe to use it for execution in the same
world. The unknown code we consider will be assumed to only have access to safe
values, which typically means that we assume it has access to a linking table
with a malloc capability and otherwise consists of instructions. This allows us
to use the FTLR seeing as malloc is a safe value (cf.
Lemma~\ref{lem:malloc-in-vr}) and instructions are really just integers and they
are always safe values. Here, it is important to remember that the FTLR gives us
that the capability for the untrusted code is safe w.r.t. some world, so even
though we described the usual assumption on the unknown code's memory, it is the
semantic model of the memory, i.e.\ the world, that the FTLR considers. In order
to argue that a specific configuration is safe, we need to argue that the
configuration is safe with respect to the world. That is, we need to argue that
the memory satisfies the world and the register-file is in the
$\stdrr$-relation. For instance, in the case where the untrusted code assumes
control first, we will use the FTLR to show that the capability for the unknown
code can be used for execution. The unknown code will get access to some known,
trusted code through a capability in the initial register-file. Because the
capability is in the register-file, we will have to argue that it is safe which
we do using Lemma~\ref{lem:anti-red-obs}.

@@

In the common case \texttt{scall} is used to invoke untrusted code, and we will
only have very general assumptions about that code, in which case we will use
the FTLR (Theorem~\ref{thm:ftlr}) to argue
Assumption~\ref{item:scall:hyp-callee}, using what we know about the values it
has access to (e.g.\ linking table, malloc capability etc.). By assumption in
Hyp-Callee, the memory is safe, so it suffices to show that the register-file
contains safe values, which amounts to showing that the arguments in the call
are safe.

@@

\begin{lemma}
  \label{lem:correctness-g1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\rwx,\glob),\dots) & c_{g1} & \;\defeq ((\entry,\glob),\dots)
    \end{array}
\]
  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1}.
  Then if $
  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),$
  then $m'(\flag) = 0$.
\end{lemma}
\begin{toplas}
  \begin{proof}[Proof sketch]
    Define a world $W_1$ with the following regions
    \begin{itemize}
    \item A malloc region, $\iota_{\malloc,0}$.
    \item A permanent region for the linking table that only accepts $\ms_\link$ and requires everything to be in $\stdvr$.
    \item A stack region, $\iota^{\pwl}_{b_\stk,e_\stk}$.
    \item A adversary region, $\iota^{\nwl,p}_{b_\adv,e_\stk}$
    \item A permanent static region for the flag and \texttt{g1}, i.e.\ a region that only
      accepts $\ms_\flag$ and $\ms_{g1}$.
    \end{itemize}
    If we can show
    \begin{equation}
      (\reg_0[\pcreg \mapsto c_\adv,r_\stk\mapsto c_\stk,r_1\mapsto c_{g1}],\ms_\malloc \uplus \ms_\link \uplus \ms_\stk \uplus \ms_\adv\uplus \ms_\flag  \uplus \ms_{g1}) \in \observations(W_1),\label{eq:g1:main-O-res}
  \end{equation}
  \footnote{We ignore step indexes in this proof sketch.}then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$ which in particular means that it respects the permanent static region that governs the assertion flag.

>>>>

\begin{figure}[b]
  \begin{tikzpicture}[scale=.7, every node={scale=.7}]
    \draw[fill=gray!20,draw=none] (0,4) rectangle (6,3.5);
    \draw[fill=gray!20,draw=none] (0,0.5) rectangle (6,1);
    \draw[draw=gray!80] (0,0.5) -- (0,4);
    \draw[draw=gray!80] (6,0.5) -- (6,4);
    \draw (0,1) rectangle (6,1.5) node[pos=.5] {\footnotesize previous stack pointer};
    \draw (0,1.5) rectangle (6,2) node[pos=.5] {\footnotesize previous program counter};
    \draw (0,2) rectangle (6,3.5) node[pos=.5] {\footnotesize restore instructions};
    \draw (-3,3.25) node (rp) {return pointer};
    \draw[->] (rp.east) -- node[above] {$\entry$}(-.2,3.25);
  \end{tikzpicture}

@@

\begin{toplas}
  In the future world relations, we must remember all region names in order to
  keep them extensional which is why we need to use masks. It does, however, not
  matter what kind of region is used to mask as it is the region name we must
  retain. This is why we allow revoked regions to be replaced by other
  regions in the two future world relations. This may seem unnecessary as the
  future world relation allows new regions to be added, but in private future
  worlds, for technical reasons, we need to be able to reinstate old regions
  with the same region names as they had in a past world. Further, we do not
  want public future worlds to be able to prevent reinstation in private future
  worlds. Private future worlds cannot revoke permanent regions, so we cannot
  allow public future worlds to us permanent regions as a mask for a revoked
  region.
\end{toplas}

@@

\begin{figure}[htb]
  \centering
  \begin{align*}
   \readCond{}(\gl)(W) &=  \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{array}{l}
       \exists r \in \var{localityReg}(g,W) \ldotp \\
\quad   \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{array}
    \right\}\\
   \writeCond{}(\iota,\gl)(W) &=  \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & W(r) \text{ is address-stratified and} \\[-0.2cm]
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
   \execCond{}(\gl)(W) &= 
    \left\{
      \npairP{(\var{P}, \start,\addrend)}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start',\addrend'] \subseteq [\start,\addrend], \perm \in \var{P} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start',\addrend',\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
   \entryCond{}(\gl)(W) &= 
    \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{array}{l}
      \forall n' < n \ldotp \forall W' \future W \ldotp\\
      \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{array}
    \right\} \\
   \text{where } \gl = \local \Rightarrow \future = \futurewk \text{ and } \gl = \glob \Rightarrow \future = \futurestr \span
   \end{align*}
\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
\begin{figure}[htb]
  \begin{align*}
  \iota^\pwl_A, \iota^\nwl_A :{} & \Regions \\
  \iota^\pwl_A \defeq{} & (\temp,1,=,=,H^\pwl_A) \\
  \iota^\nwl_A \defeq{} & (\temp,1,=,=,H^\nwl_A) \\
  \iota^{\nwl,p}_A \defeq{} & (\perma,1,=,=,H^\nwl_A)\\
    H^\pwl_A :{} & \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
  H^\pwl_A \; s \; \hat{W} \defeq{} & \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = A \land \\
      &\forall \addr \in A \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
    \end{aligned}
        \right\} \union \{\npair[0]{\ms} \} \\
    H^\nwl_A :{} & \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
  H^\nwl_A \; s \;\hat{W} \defeq{} & \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = A \land \\
      &\forall \addr \in A \ldotp \\
      &\quad \nonlocal{\ms(\addr)} \land\\ 
      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
    \end{aligned}
        \right\} \union \{\npair[0]{\ms} \}
\end{align*}

@@

\begin{toplas}
  In \sectionname~\ref{sec:stack-and-return-pointer}, we require capabilities with write-local permission to be local.
  This would indicate that our logical relation should imply the same, namely that capabilities with write-local permission are local:
  \begin{lemma}[Stack capabilities are local]~
    \label{lem:wl-local}
    \begin{itemize}
    \item If $\memSat{\ms}{W}$ and $\npair{\stdcap} \in \stdvr(W)$ and $\start
      \leq \addrend$ and $\perm \in \{\rwlx,\rwl\}$, then $\gl = \local$.
    \end{itemize}
  \end{lemma}
  In our definition of worlds, nothing prevents a world from having regions that are overlapping.
  This might seem like an issue the $\stdvr$-relation as it allows different permission-based requirements to be satisfied by different regions.
  In practice, it is not an issue as we will always have a memory satisfaction assumption which doubles as a well-formedness condition on the world as it prevents the regions from overlapping.
  The memory satisfaction assumption in Lemma~\ref{lem:wl-local} is there to ensure a well-formed world.
\end{toplas}

\subsection{Safety of the Capability Machine}
With the logical relation defined, we can now state the fundamental theorem of
our logical relation: a strong theorem that formalizes the guarantees offered by
the capability machine. Essentially, it says a capability that only grants safe
authority is capability safe as a program counter.
\begin{theorem}[Fundamental Theorem]
  \label{thm:ftlr}
  If one of the following holds:
  \begin{align*}
      & \bullet
        \begin{aligned}[t]
        &\perm = \exec \text{ and }\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwx \text{ and } \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwlx \text{ and }\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
      \end{aligned}
  \end{align*}
  then $\npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)$
\end{theorem}
\begin{toplas}
  \begin{proof}[Proof sketch]
    Induction over $n$. By definition of $\stder(W)$, show
    \[
      \npair[n']{(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms)}
      \in \observations(W)
    \]
    assuming $n' \leq n$, $\npair[n']{\reg} \in \stdrr(W)$, and
    $\heapSat[\ms]{n'}{W}$. By definition of $\observations$ let $\ms_f$,
    $\mem'$, and $i \leq n'$ be given and for
    $\Phi = (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms \uplus
    \ms_f)$ assume $\Phi \step[i]
    (\halted,\mem')$ and show there exists $W' \futurestr W$ that part of
    $\mem'$ satisfies. First observe that $i \neq 0$ as $\Phi$ is a non-halted
    configuration, so $\Phi$ takes at least one step, i.e.\ $\Phi \step
    \Phi' \step[i-1] (\halted,\mem')$.

@@

In addition to these syntactic requirements, we also specify standard
expectations for how malloc behaves. Intuitively, we require that when malloc is
invoked with a length argument, then it will return a capability for a fresh
piece of memory of that size. It should be fresh in the sense that malloc has
not already given out a capability for any part of that memory before and will
not do so in the future. Also, when invoked with a nonsensical length argument
such as a negative integer or a capability, malloc should simply fail. However,
formulating these requirements is harder than one might expect. The problem is
that a realistic implementation of malloc needs to rely on internal state that
changes after every invocation and relies on invariants on that state. We can
only expect that malloc behaves according to its specification, if its internal
state satisfies the current state of the invariants in an executing system. To
express this, we will use the semantic model defined in
Section~\ref{sec:logical-relation}.

@@

\begin{figure}[tb]
    \centering
\begin{lstlisting}
 $i_1$ = move r_t1 pc
 $i_2$ = lea r_t1 $\var{ofc}$
 $i_3$ = load r_stk r_t1
 $i_4$ = pop pc
\end{lstlisting}
    \caption{The restoration code used in \texttt{scall}. The variable
      $\var{ofc}$ is 5 which is the offset to the address where the old stack
      pointer is stored on the stack.}
    \label{fig:scall-rest-code}
\end{figure}

\begin{figure}[htb]
  \begin{toplas}
    \centering
\begin{lstlisting}
// push private registers to the stack
  push $r_{\var{priv},1}$
  ...
  push $r_{\var{priv},n}$
// push restoration code to the stack
  push $\encode(i_1)$
  push $\encode(i_2)$
  push $\encode(i_3)$
  push $\encode(i_4)$
// push old pc to the stack
  move r_t1 pc
  lea r_t1 $\var{ofc}$ 
  push r_t1
// push stack pointer to the stack
  push r_stk
// set up protected return pointer
  move r_0 r_stk
  lea r_0 $\var{ofc_\var{rec}}$ // -5 is the offset to the first instruction of
the recovery code
  restrict r_0 $\encodePermPair((\local,\entry))$
// restrict stack capability to unused part
  geta r_t1 r_stk
  plus r_t1 r_t1 1
  getb r_t2 r_stk
  subseg r_stk r_t1 r_t2
// clear unused part of the stack
  mclear r_stk
// clear non-argument registers
  rclear $R$
  jmp r
after:
// pop the restore code
  pop r_t1
  pop r_t1
  pop r_t1
  pop r_t1
// pop the private state into approriate registers
  pop $r_{\var{priv},1}$
  ...
  pop $r_{\var{priv},n}$
\end{lstlisting}
    \caption{Implementation of \texttt{scall} $r(r_{\var{args},1},\dots,
      r_{\var{args},m},r_{\var{priv},1},\dots, r_{\var{priv},n})$. The
      restoration code is given in \figurename~\ref{fig:scall-rest-code}. The
      variable $\var{ofc}$ is the offset to the label after, and
      $\var{ofc_\var{rec}} = -5$ which is the offset to the first instruction of
      the activation record. The set $R = \RegName \setminus
      \{\texttt{pc},\texttt{r\_stk},\texttt{r\_0},r,r_{\var{args},1},\dots,r_{\var{args},m}\}$.}
    \label{fig:scall-impl}
  \end{toplas}
\end{figure}

@@

% Anti-reduction lemma
Our capability machine only allows the final memory of an execution to be observed, so naturally the correctness lemmas we prove in \sectionname~\ref{sec:examples} are statements about the memory in the halted configuration.
In order to prove something about some part of the final memory, we establish an invariant for it as a region in a world and show that the initial configuration is in the $\observations$-relation w.r.t.\ that world.
The invariants we are interested in must be permanent regions so that they are not revoked during execution.
The $\observations$-relation says that if the initial configuration halts successfully, then the memory in the halting state must still respect a private future world of the initial world.
It is, however, bothersome and error prone to try to argue about the entire execution in one go.
Instead we want to reason modularly, in the sense that we only want to reason about parts of the execution at a time.
To this end, we prove an anti-reduction lemma that essentially says that if we can show that an initial configuration, $\Phi$, steps to a configuration $\Phi'$ and that $\Phi'$ is in the $\observations$-relation, then $\Phi$ is also in the observation relation.
\begin{lemma}[Anti-reduction for $\observations$]
\label{lem:anti-red-obs}
  \begin{align*}
    & \forall n, n',i,\reg, \reg', \ms, \ms', \ms_r, W, W' \ldotp\\
    & \quad n' \geq n - i \land W' \futurestr W \land \\
    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \\
    & \quad \npair[n']{(\reg',\ms')} \in \observations(W') \\
    & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
  \end{align*}
\end{lemma}

% FTLR
The anti-reduction works when we reason about the execution of known code,
because we can execute it. When we want to reason about unknown code, the
anti-reduction lemma does not apply because we do not know what instructions are
being executed. This is where the FTLR (Theorem~\ref{thm:ftlr}) comes into play.
As a reminder, the FTLR says that if a capability only has access to safe values
with respect to a world, then it is safe to use it for execution in the same
world. The unknown code we consider will be assumed to only have access to safe
values, which typically means that we assume it has access to a linking table
with a malloc capability and otherwise consists of instructions. This allows us
to use the FTLR seeing as malloc is a safe value (cf.
Lemma~\ref{lem:malloc-in-vr}) and instructions are really just integers and they
are always safe values. Here, it is important to remember that the FTLR gives us
that the capability for the untrusted code is safe w.r.t. some world, so even
though we described the usual assumption on the unknown code's memory, it is the
semantic model of the memory, i.e.\ the world, that the FTLR considers. In order
to argue that a specific configuration is safe, we need to argue that the
configuration is safe with respect to the world. That is, we need to argue that
the memory satisfies the world and the register-file is in the
$\stdrr$-relation. For instance, in the case where the untrusted code assumes
control first, we will use the FTLR to show that the capability for the unknown
code can be used for execution. The unknown code will get access to some known,
trusted code through a capability in the initial register-file. Because the
capability is in the register-file, we will have to argue that it is safe which
we do using Lemma~\ref{lem:anti-red-obs}.

@@

In the common case, \texttt{scall} is used to invoke untrusted code, and we will
only have very general assumptions about that code, in which case we will use
the FTLR (Theorem~\ref{thm:ftlr}) to argue
Assumption~\ref{item:scall:hyp-callee}, using what we know about the values it
has access to (e.g.\ linking table, malloc capability etc.). By assumption in
Hyp-Callee, the memory is safe, so it suffices to show that the register-file
contains safe values, which amounts to showing that the arguments in the call
are safe.

@@

\begin{lemma}
  \label{lem:correctness-g1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\rwx,\glob),\dots) & c_{g1} & \;\defeq ((\entry,\glob),\dots)
    \end{array}
\]
  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1}.
  Then if $
  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),$
  then $m'(\flag) = 0$.
\end{lemma}
\begin{toplas}
  \begin{proof}[Proof sketch]
    Define a world $W_1$ with the following regions
    \begin{itemize}
    \item A malloc region, $\iota_{\malloc,0}$.
    \item A permanent region for the linking table that only accepts $\ms_\link$ and requires everything to be in $\stdvr$.
    \item A stack region, $\iota^{\pwl}_{b_\stk,e_\stk}$.
    \item An adversary region, $\iota^{\nwl,p}_{b_\adv,e_\stk}$
    \item A permanent static region for the flag and \texttt{g1}, i.e.\ a region that only
      accepts $\ms_\flag$ and $\ms_{g1}$.
    \end{itemize}
    If we can show
    \begin{equation}
      (\reg_0[\pcreg \mapsto c_\adv,r_\stk\mapsto c_\stk,r_1\mapsto c_{g1}],\ms_\malloc \uplus \ms_\link \uplus \ms_\stk \uplus \ms_\adv\uplus \ms_\flag  \uplus \ms_{g1}) \in \observations(W_1),\label{eq:g1:main-O-res}
  \end{equation}
  \footnote{We ignore step indexes in this proof sketch.}then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$ which in particular means that it respects the permanent static region that governs the assertion flag.

  As described in \sectionname~\ref{sec:reasoning}, we use the FTLR (Theorem~\ref{thm:ftlr}) to reason about unknown code, so we use it to reason about $c_\adv$.
  With the standard region $\iota^{nwl,p}$ chosen for the adversary, it is easy to show that the $\readCond{}$ and $\writeCond{}$ holds for $c_\adv$ which gives us $c_\adv \in \stder(W)$ by the FTLR.
  In order to get \eqref{eq:g1:main-O-res}, we need to show that (a) the memory satisfies the world and (b) the register file is in the $\stdrr(W)$.
  We have defined the world, so there (almost) is a one-to-one correspondence between region and memory segment in the initial configuration, so (a) is easily shown.
  In order to show (b), we need to show $c_\stk \in \stdvr(W)$ and $c_{g1} \in \stdvr(W)$.
  The former follows easily from the choice of a $\iota^{pwl}$-region for the stack in $W$.
  In order to argue the latter, we basically have to argue that $c_{g1}$, the closure generator, respects the world $W$.
  This amounts to showing that the closures generated by \texttt{g1} also respect the invariants of $W$.
  We reason about the local variable $x$ in the closure in the same way \citet{Dreyer:jfp12} does.
  We ignore $x$ in the remainder of this proof sketch to focus on the parts of the proof related to the setting of the capability machine.
